|

### 2023-07-21 journal/response.Rmd

> **We appreciate the reviewer's feedback and motivation for a regularized adjusted plus-minus (RAPM) style approach for modeling STRAIN. Since our response variable is a pass rusher's average STRAIN for a play, we are not necessarily interested in estimating effects for every player since the interpretation of the coefficient would be at a player-play level. However, we have updated text in Section 5 about modeling an aggregate STRAIN using a RAPM style approach similar to Sabin (2021). Since the primary purpose of this manuscript was to introduce STRAIN and its usefulness in measuring pass rush ability, we leave the RAPM approach for future work in a separate manuscript.**

# Associate Editor

1. The first comment that warrants serious consideration is whether the average strain, across an entire play, is the best way to summarize a pash rushers effect on a particular play. Furthermore, and as just one example, would this summary consistently capture the pash rusher effect across plays of different length?

> **In addition to the average STRAIN, we also considered other summaries, namely, minimum STRAIN and maximum STRAIN. We chose average STRAIN since it is a highly stable quantity over time, as demonstrated in Figure 6. On the other hand, we also compute the correlations in minimum and maximum STRAIN between the first and last four weeks of the 2021 NFL season, and they are both much lower compared to average STRAIN. 
Further, we recognize the limitation that our analysis did not account for the fact that play length can differ. This problem can be addressed with a time-to-event model. Although we agree that this is an important consideration, since the main goal of this paper is to introduce STRAIN as a metric for evaluating pass rushers in football, we leave this for future work.

### 2023-07-21 journal/response.Rmd

> **We appreciate the reviewer's feedback and motivation for a regularized adjusted plus-minus (RAPM) style approach for modeling STRAIN. Since our response variable is a pass rusher's average STRAIN for a play, we are not necessarily interested in estimating effects for every player since the interpretation of the coefficient would be at a player-play level. However, we have updated text in Section 5 about modeling an aggregate STRAIN using a RAPM style approach similar to Sabin (2021). Since the primary purpose of this manuscript was to introduce STRAIN and its usefulness in measuring pass rush ability, we leave the RAPM approach for future work in a separate manuscript.**

# Associate Editor

1. The first comment that warrants serious consideration is whether the average strain, across an entire play, is the best way to summarize a pash rushers effect on a particular play. Furthermore, and as just one example, would this summary consistently capture the pash rusher effect across plays of different length?

> **In addition to the average STRAIN, we also considered other summaries, namely, minimum STRAIN and maximum STRAIN. We chose average STRAIN since it is a highly stable quantity over time, as demonstrated in Figure 6. On the other hand, we also compute the correlations in minimum and maximum STRAIN between the first and last four weeks of the 2021 NFL season, and they are both much lower compared to average STRAIN. 
Further, we recognize the limitation that our analysis did not account for the fact that play length can differ. This problem can be addressed with a time-to-event model. Although we agree that this is an important consideration, since the main goal of this paper is to introduce STRAIN as a metric for evaluating pass rushers in football, we leave this for future work.**

### 2023-07-22 journal/response.Rmd

> **We appreciate the reviewer's feedback and motivation for a regularized adjusted plus-minus (RAPM) style approach for modeling STRAIN. Since our response variable is a pass rusher's average STRAIN for a play, we are not necessarily interested in estimating effects for every player since the interpretation of the coefficient would be at a player-play level. However, we have updated text in Section 5 about modeling an aggregate STRAIN using a RAPM style approach similar to Sabin (2021). Since the primary purpose of this manuscript was to introduce STRAIN and its usefulness in measuring pass rush ability, we leave the RAPM approach for future work in a separate manuscript.**

# Associate Editor

1. The first comment that warrants serious consideration is whether the average strain, across an entire play, is the best way to summarize a pash rushers effect on a particular play. Furthermore, and as just one example, would this summary consistently capture the pash rusher effect across plays of different length?

> **In addition to the average STRAIN, we also considered other summaries, namely, minimum STRAIN and maximum STRAIN. We chose average STRAIN since it is a highly stable quantity over time, as demonstrated in Figure 6. On the other hand, we also compute the correlations in minimum and maximum STRAIN between the first and last four weeks of the 2021 NFL season, and they are both much lower compared to average STRAIN. 
Further, we recognize the limitation that our analysis did not account for the fact that play length can differ. Although we agree that this is an important consideration, since the main goal of this paper is to introduce STRAIN as a metric for evaluating pass rushers in football, we leave consideration of time-to-event modeling for future work.**

### 2023-07-22 journal/response.Rmd

---
title: "Response to Reviewers"
output: pdf_document
fontsize: 11pt
header-includes:
 \usepackage{mathptmx}
---

@@

> **We have added more descriptions of the fixed effects in our multilevel model. See the last paragraph of Section 3.2 in our revised paper.**

# Reviewer 2

* Page 3: It seems the author is defining anyone who lines up on the line of scrimmage outside a DE. This is not always agreed upon, I would recommend the authors make how they're defining that clear. Is it per play? Or is one player defined as always being a DE?

> **We thank the reviewer for the recommendation. Here, we are showing where a DE typically (not always) lines up, with regards to the two defensive schemes in Figure 1 (of the original manuscript). In the revised version, this part has been moved to the supplement; see section A. We have added clarification about the positions and it now reads "Note that apart from the formations shown in Figure A.1 [of the supplement], defensive linemen can have the flexibility to line up differently. For example, a defensive end, depending on their opposing matchup, may not necessarily be positioned toward the outside of the line of scrimmage."**

@@

> **We appreciate the reviewer's feedback and motivation for a regularized adjusted plus-minus (RAPM) style approach for modeling STRAIN. Since our response variable is a pass rusher's average STRAIN for a play, we are not necessarily interested in estimating effects for every player since the interpretation of the coefficient would be at a player-play level. However, we have updated text in Section 5 about modeling an aggregate STRAIN using a RAPM style approach similar to Sabin (2021). Since the primary purpose of this manuscript was to introduce STRAIN and its usefulness in measuring pass rush ability, we leave the RAPM approach for future work in a separate manuscript.**

# Associate Editor

1. The first comment that warrants serious consideration is whether the average strain, across an entire play, is the best way to summarize a pash rushers effect on a particular play. Furthermore, and as just one example, would this summary consistently capture the pash rusher effect across plays of different length?

> **In addition to the average STRAIN, we also considered other summaries, namely, minimum STRAIN and maximum STRAIN. We chose average STRAIN since it is a highly stable quantity over time, as demonstrated in Figure 6. On the other hand, we also compute the correlations in minimum and maximum STRAIN between the first and last four weeks of the 2021 NFL season, and they are both much lower compared to average STRAIN. 
Further, we recognize the limitation that our analysis did not account for the fact that play length can differ. Although we agree that this is an important consideration, which requires time-to-event modeling, since the main goal of this paper is to introduce STRAIN as a metric for evaluating pass rushers in football, we leave this problem for future work.**

### 2023-07-22 journal/01-introduction.Rmd
last revision before submission

Our focus of this manuscript is specifically on measuring the performance of defensive linemen in the NFL.
Using data made available in the Big Data Bowl 2023, we present a novel approach to measure the performance of pass rushers.
Relative to other aspects of American football, such as quarterback evaluation [@burke2019deepqb; @reyers2021quarterback], the literature on evaluating pass rushers is scarce.
Below, we provide a brief overview of existing pass rush metrics.

@@

<!-- Table \ref{tab:metrics} gives a summary of existing football metrics for pass rush. -->
<!-- We now highlight what these quantities describe as well as their limitations. -->

@@

In addition, the sum of sacks, hits, and hurries is often defined as pressures.
This is better than the individual counts to some extent, but suffers from problems of subjectivity (e.g., whether there is an actual hurry or not).
Pro Football Focus (PFF) defines a metric called pass-rush productivity, which is a minor modification from the aforementioned pressures metric.<!-- (see Table \ref{tab:metrics}). -->
In particular, pass-rush productivity gives twice as much weight to a sack relative to hurries and hits, which is a small upgrade to pressures.
However, the choice of weights is ad-hoc and still only considers binary outcomes, similar to the shortcomings of previous metrics.

>>>>

Our focus of this manuscript is specifically on measuring the performance of NFL defensive linemen (see Section A of the Supplement for more information on the defensive positions in football).
Using data made available in the Big Data Bowl 2023, we present a novel approach to measure the performance of pass rushers.
Relative to other aspects of American football, such as quarterback evaluation [@burke2019deepqb; @reyers2021quarterback], the literature on evaluating pass rushers is scarce.
Below, we provide a brief overview of existing pass rush metrics.

@@

Table B.1 of the Supplement gives a summary of existing football metrics for pass rush.
We now highlight what these quantities describe as well as their limitations.

@@

In addition, the sum of sacks, hits, and hurries is often defined as pressures.
This is better than the individual counts to some extent, but suffers from problems of subjectivity (e.g., whether there is an actual hurry or not).
Pro Football Focus (PFF) defines a metric called pass-rush productivity, which is a minor modification from the aforementioned pressures metric.
In particular, pass-rush productivity gives twice as much weight to a sack relative to hurries and hits, which is a small upgrade to pressures.
However, the choice of weights is ad-hoc and still only considers binary outcomes, similar to the shortcomings of previous metrics.

### 2023-07-22 journal/02-data.Rmd
last revision before submission

In the forthcoming analysis, we rely on the data from the NFL Big Data Bowl 2023 provided by the NFL Next Gen Stats tracking system.
The data corresponds to 8,557 passing plays across 122 games in the first eight weeks of the 2021 NFL regular season. 
For each play, we have information on the on-field location, speed, angle, direction, and orientation of each player on the field and the football at a rate of 10 Hz (i.e., 10 measurements per second), along with event annotations for each frame such as ball snap, pass forward, and quarterback sack, to name a few.
Table \ref{tab:track} displays a tracking data example for a [play](https://www.raiders.com/video/de-maxx-crosby-sacks-qb-teddy-bridgewater-for-a-loss-of-6-yards-nfl) from the 2021 NFL regular season week six matchup between the Las Vegas Raiders and Denver Broncos, which ends with Broncos quarterback Teddy Bridgewater getting sacked by Raiders defensive end Maxx Crosby.

@@

@@

\begin{table}
\caption{Example of tracking data for a play during the Las Vegas Raiders versus Denver Broncos NFL game on October 17, 2021. The data shown here are for Raiders defensive end Maxx Crosby, and the frames included are between the ball snap and when the sack by Crosby is recorded. \label{tab:track}}
\centering
\begin{tabular}{ccccccccc}
\hline
frameId & x & y & s & a & dis & o & dir & event \\ 
\hline
7 & 67.68 & 29.89 & 0.34 & 1.57 & 0.04 & 124.86 & 88.21 & ball\_snap \\ 
8 & 67.76 & 29.89 & 0.69 & 2.13 & 0.08 & 124.07 & 89.59 & None \\ 
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\ 
50 & 73.67 & 25.06 & 4.19 & 2.62 & 0.42 & 134.21 & 125.26 & qb\_sack \\ 
\hline
\end{tabular}
\end{table}

@@

Along with the tracking information, the Big Data Bowl 2023 includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
We also leverage this scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

>>>>

In the forthcoming analysis, we rely on the data from the NFL Big Data Bowl 2023 provided by the NFL Next Gen Stats tracking system.
The data corresponds to 8,557 passing plays across 122 games in the first eight weeks of the 2021 NFL regular season. 
For each play, we have information on the on-field location, speed, angle, direction, and orientation of each player on the field and the football at a rate of 10 Hz (i.e., 10 measurements per second), along with event annotations for each frame such as ball snap, pass forward, and quarterback sack, to name a few.

@@

Throughout this paper, we use a [play](https://www.raiders.com/video/de-maxx-crosby-sacks-qb-teddy-bridgewater-for-a-loss-of-6-yards-nfl) from the 2021 NFL regular season week six matchup between the Las Vegas Raiders and Denver Broncos as a case study for examining pass rush in football.
Table C.1 of the Supplement displays a tracking data example for the aforementioned play, which ends with Broncos quarterback Teddy Bridgewater getting sacked by Raiders defensive end Maxx Crosby.

@@

@@

Along with the tracking information, the Big Data Bowl 2023 includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
For context, Figure D.1 of the Supplement displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e., 4-man rush formation) as the most common value.
We also leverage this scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

### 2023-07-22 journal/03-methods.Rmd
last revision before submission

Moreover, since we observe STRAIN at every tenth of a second within each play, we can then compute the average STRAIN across all frames played for every pass rusher.
Formally, the average STRAIN, denoted by $\overline{\text{STRAIN}}$, for pass rusher $j$ involved in $n_j$ total plays across $\sum_{i\in Z_j} T_i$ total frames, where $Z_j$ is the set of all plays with pass rusher $j$'s involvement, is
\begin{equation*}
\overline{\text{STRAIN}}_{j} = \frac{1}{\sum_{i\in Z_j} T_i} \sum_{i\in Z_j} \sum_{t=1}^{T_i} \widehat{\text{STRAIN}}_{ij}(t).
\end{equation*}
This can be helpful for metric assessment, as we use average STRAIN to examine different statistical properties of our measure in Section \ref{sec:statprop}.

@@

In detail, we consider a normal distribution to shrink the random intercepts for each player and team toward their respective group means.
This is a useful property since we do not observe the same number of plays for each player. 
For the team effects, this provides us with the average defense and offense team-level effects on a pass rusher's STRAIN.
Due to the nested nature of players on teams, the individual pass rusher and blocker random intercepts reflect the respective player's effect relative to their team effects. 
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language].

>>>>

Moreover, since we observe STRAIN at every tenth of a second within each play, we can then compute the average STRAIN across all frames played for every pass rusher.
Formally, the average STRAIN, denoted by $\overline{\text{STRAIN}}$, for pass rusher $j$ involved in $n_j$ total plays across $\sum_{i\in Z_j} T_i$ total frames, where $Z_j$ is the set of all plays with pass rusher $j$'s involvement, is
\begin{equation*}
\overline{\text{STRAIN}}_{j} = \frac{1}{\sum_{i\in Z_j} T_i} \sum_{i\in Z_j} \sum_{t=1}^{T_i} \widehat{\text{STRAIN}}_{ij}(t).
\end{equation*}
This can be helpful in determining the most effective pass rushers; see Section E of the Supplement for our pass rusher evaluation based on average STRAIN.
We also use average STRAIN to assess different statistical properties of our metric in Section \ref{sec:statprop}.

@@

In detail, we consider a normal distribution to shrink the random intercepts for each player and team toward their respective group means.
This is a useful property since we do not observe the same number of plays for each player. 
For the team effects, this provides us with the average defense and offense team-level effects on a pass rusher's STRAIN.
Due to the nested nature of players on teams, the individual pass rusher and blocker random intercepts reflect the respective player's effect relative to their team effects.
Furthermore, our proposed model demonstrates better out-of-sample predictive performance than a model without random intercepts for teams, as detailed in Section F of the Supplement.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language].

### 2023-07-22 journal/04-results.Rmd
minor edits; ready for review
last revision before submission

To illustrate our proposed metric STRAIN for pass rush evaluation, we use the same play from the Las Vegas Raiders and Denver Broncos game as mentioned in Section \ref{sec:data}.
Figure \ref{fig:fig_features_combined} (left) is a line graph showing how Crosby's distance from the quarterback, velocity, and STRAIN change continuously throughout the play.
We observe that for the first two seconds, Crosby is being blocked by a Denver offensive lineman and unable to get close to the quarterback, hence the corresponding STRAIN values are virtually zero. 
Suddenly at around three seconds after the snap, the STRAIN for Crosby starts to increase after the Raiders defensive end is freed up and charges towards the QB. 
At 4 seconds after the snap, Crosby's STRAIN is 2.30, which means at his current moving rate, it will take Crosby about 0.43 (1/2.30) seconds to make the distance between him and the quarterback 0 (i.e. essentially sack the quarterback).
This matches well with the final outcome of the play, as the sack takes place at the very last frame (4.4 seconds) where the estimated STRAIN for Crosby reaches its peak at 3.96.

@@

The results of fitting the multilevel model described in Section \ref{sec:multilevel} is displayed in Table \ref{tab:fixedeff}.
First, we investigate the fixed-effects terms of this model.
It appears that that the average STRAIN decreases by 0.7366 (95% CI [-0.7875, -0.6858]) for every additional blocker in the offensive unit, after accounting for other covariates.
In other words, NFL pass rushers tend to generate more pressure when facing fewer number of blockers, which makes intuitive sense.

@@

\begin{table}
\caption{Fixed-effects coefficient estimates for the multilevel model for average STRAIN. Note that the reference down level is first down, the reference pass rusher level is defensive end (DE), denoted R:DE; the reference pass blocker level is tackle (T), denoted B:T. \label{tab:fixedeff}}
\centering
\begin{tabular}{lrrrr}
\hline
& estimate & se & $t$-statistic & $p$-value \\ 
\hline
Intercept & 1.8004 & 0.0991 & 18.1723 & 0.0000 \\ 
Number of blockers & -0.7366 & 0.0259 & -28.4153 & 0.0000 \\
Yards to go & 0.0491 & 0.0033 & 14.9013 & 0.0000 \\
Current yardline & 0.0001 & 0.0005 & 0.1844 & 0.8537 \\
$I_{\text{\{2nd down\}}}$ & 0.5143 & 0.0290 & 17.7439 & 0.0000 \\
$I_{\text{\{3rd down\}}}$ & 0.9176 & 0.0324 & 28.3613 & 0.0000 \\
$I_{\text{\{4th down\}}}$ & 0.4649 & 0.0767 & 6.0637 & 0.0000 \\
$I_{\text{\{2pt conversion\}}}$ & -0.1384 & 0.1797 & -0.7699 & 0.4413 \\
$I_{\text{\{R:DT\}}}$ & -0.7166 & 0.0678 & -10.5658 & 0.0000 \\
$I_{\text{\{R:interior\}}}$ & 0.5611 & 0.0999 & 5.6143 & 0.0000 \\
$I_{\text{\{R:NT\}}}$ & -0.9426 & 0.1021 & -9.2344 & 0.0000 \\
$I_{\text{\{R:OLB\}}}$ & 0.5574 & 0.0726 & 7.6742 & 0.0000 \\
$I_{\text{\{R:secondary\}}}$ & 1.4252 & 0.1114 & 12.7899 & 0.0000 \\
$I_{\text{\{B:C\}}}$ & -0.2983 & 0.0594 & -5.0237 & 0.0000 \\
$I_{\text{\{B:G\}}}$ & -0.1546 & 0.0471 & -3.2847 & 0.0000 \\ 
$I_{\text{\{B:other\}}}$ & -0.1260 & 0.0646 & -1.9508 & 0.0514 \\
\hline
\end{tabular}
\end{table}

>>>>

To illustrate our proposed metric STRAIN for pass rush evaluation, we use the example play from the Las Vegas Raiders and Denver Broncos game as mentioned in Section \ref{sec:data}.
Figure \ref{fig:fig_features_combined} (left) is a line graph showing how Crosby's distance from the quarterback, velocity, and STRAIN change continuously throughout the play.
In addition, Figure C.1 of the Supplement presents the locations of every Las Vegas and Denver player
on the field from this play at 1, 2, 3 and 4 seconds after the ball snap, with the point size for each Las Vegas defender corresponding to the estimated STRAIN at the selected frames.

We observe that for the first two seconds, Crosby is being blocked by a Denver offensive lineman and unable to get close to the quarterback, hence the corresponding STRAIN values are virtually zero. 
Suddenly at around three seconds after the snap, the STRAIN for Crosby starts to increase after the Raiders defensive end is freed up and charges towards the QB. 
At 4 seconds after the snap, Crosby's STRAIN is 2.30, which means at his current moving rate, it will take Crosby about 0.43 (1/2.30) seconds to make the distance between him and the quarterback 0 (i.e. essentially sack the quarterback).
This matches well with the final outcome of the play, as the sack takes place at the very last frame (4.4 seconds) where the estimated STRAIN for Crosby reaches its peak at 3.96.

@@

The results of fitting the multilevel model described in Section \ref{sec:multilevel} is displayed in Table \ref{tab:fixedeff}.
First, we investigate the fixed effects terms of this model.
It appears that that the average STRAIN decreases by 0.7366 (95% CI [-0.7875, -0.6858]) for every additional blocker in the offensive unit, after accounting for other covariates.
In other words, NFL pass rushers tend to generate more pressure when facing fewer number of blockers, which makes intuitive sense.

@@

\begin{table}
\caption{Fixed effects coefficient estimates for the multilevel model for average STRAIN. Note that the reference down level is first down, the reference pass rusher level is defensive end (DE), denoted R:DE; the reference pass blocker level is tackle (T), denoted B:T. \label{tab:fixedeff}}
\centering
\begin{tabular}{lrrrr}
\hline
& estimate & se & $t$-statistic & $p$-value \\ 
\hline
Intercept & 1.8004 & 0.0991 & 18.1723 & 0.0000 \\ 
Number of blockers & -0.7366 & 0.0259 & -28.4153 & 0.0000 \\
Yards to go & 0.0491 & 0.0033 & 14.9013 & 0.0000 \\
Current yardline & 0.0001 & 0.0005 & 0.1844 & 0.8537 \\
$I_{\text{\{2nd down\}}}$ & 0.5143 & 0.0290 & 17.7439 & 0.0000 \\
$I_{\text{\{3rd down\}}}$ & 0.9176 & 0.0324 & 28.3613 & 0.0000 \\
$I_{\text{\{4th down\}}}$ & 0.4649 & 0.0767 & 6.0637 & 0.0000 \\
$I_{\text{\{2pt conversion\}}}$ & -0.1384 & 0.1797 & -0.7699 & 0.4413 \\
$I_{\text{\{R:DT\}}}$ & -0.7166 & 0.0678 & -10.5658 & 0.0000 \\
$I_{\text{\{R:interior\}}}$ & 0.5611 & 0.0999 & 5.6143 & 0.0000 \\
$I_{\text{\{R:NT\}}}$ & -0.9426 & 0.1021 & -9.2344 & 0.0000 \\
$I_{\text{\{R:OLB\}}}$ & 0.5574 & 0.0726 & 7.6742 & 0.0000 \\
$I_{\text{\{R:secondary\}}}$ & 1.4252 & 0.1114 & 12.7899 & 0.0000 \\
$I_{\text{\{B:C\}}}$ & -0.2983 & 0.0594 & -5.0237 & 0.0000 \\
$I_{\text{\{B:G\}}}$ & -0.1546 & 0.0471 & -3.2847 & 0.0000 \\ 
$I_{\text{\{B:other\}}}$ & -0.1260 & 0.0646 & -1.9508 & 0.0514 \\
\hline
\end{tabular}
\end{table}

### 2023-07-22 journal/supplement.Rmd
last revision before submission

---
title: |
  | Supplement to "Here Comes the STRAIN: Analyzing Defensive Pass Rush in American Football with Player Tracking Data"
blinded: 0
bibliography: bibliography.bib
biblio-style: chicago
authors: 
- name: Quang Nguyen
  # thanks: Corresponding author's email
  affiliation: |
    | Department of Statistics & Data Science
    | Carnegie Mellon University
- name: Ronald Yurko
  affiliation: |
    | Department of Statistics & Data Science
    | Carnegie Mellon University
- name: Gregory J. Matthews
  affiliation: |
    | Department of Mathematics and Statistics
    | Center for Data Science and Consulting
    | Loyola University Chicago
output: rticles::asa_article
header-includes:
  \date{}
  \def\thesection{\Alph{section}}
---

\spacingset{1.5}

@@

Table \ref{tab:metrics} gives a summary of existing football metrics for pass rush, as discussed in Section 1.1 of the paper.

@@

@@

Figure \ref{fig:fig_field} shows snapshots of a play from the week six game between the Las Vegas Raiders and Denver Broncos in the 2021 NFL regular season.
The play outcome is a sack made by Raiders defensive end Maxx Crosby on Broncos quarterback Teddy Bridgewater.
The locations of every Las Vegas (in black, on defense) and Denver (in orange, on offense) player on the field at 1, 2, 3, and 4 seconds after the ball snap is displayed, with Crosby highlighted in blue.
Note that the point size for each Las Vegas player corresponds to their estimated STRAIN as the play progresses.
This play is illustrated with Table 1 and Figure 1 (left) of the paper.

@@

Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play.
The number of pass rusher ranges from 1 to 8, with 4 pass rushers (i.e., 4-man rush formation) as the most common value.

@@

# Comparison of Multilevel Models

## Modeling Play-Level STRAIN Without Team Effects

@@

## Model Comparison

We perform leave-one-week-out cross-validation to compare the multilevel models with and without team effects for play-level STRAIN.
For example, we train each model using data from weeks one through seven, and perform evaluation for week eight.
This procedure is repeated for every single holdout week, and we evaluate the model performance based on mean squared error (MSE).
Figure \ref{fig:fig_loocv_team_eff} shows the week-by-week holdout MSE for the aforementioned multilevel models.
We clearly see that a smaller MSE is achieved for each left-out week when accounting for team effects.
Furthermore, the model with team effects yields a lower overall holdout MSE ($4.1882$) than the one without random intercepts for teams ($4.4138$).
The out-of-sample performance justifies the inclusion of team effects in our multilevel model.

>>>>

---
title: |
  | Supplement to "Here Comes the STRAIN: Analyzing Defensive Pass Rush in American Football with Player Tracking Data"
blinded: 0
bibliography: bibliography.bib
biblio-style: chicago
authors: 
- name: Quang Nguyen
  # thanks: Corresponding author's email
  affiliation: |
    | Department of Statistics & Data Science
    | Carnegie Mellon University
- name: Ronald Yurko
  affiliation: |
    | Department of Statistics & Data Science
    | Carnegie Mellon University
- name: Gregory J. Matthews
  affiliation: |
    | Department of Mathematics and Statistics
    | Center for Data Science and Consulting
    | Loyola University Chicago
output: rticles::asa_article
header-includes:
  \date{}
  \def\thesection{\Alph{section}}
  \counterwithin{figure}{section}
  \counterwithin{table}{section}
---

\spacingset{1.45}

@@

Table \ref{tab:metrics} summarizes previous pass rush statistics in football, to accompany Section 1.1 of the paper.

@@

\newpage

@@

Table \ref{tab:track} shows an example of player tracking data for a play from the week six game between the Las Vegas Raiders and Denver Broncos in the 2021 NFL regular season.
The play outcome is a sack made by Raiders defensive end Maxx Crosby on Broncos quarterback Teddy Bridgewater.
Figure \ref{fig:fig_field} shows different snapshots of this play, where the locations of every Las Vegas (in black, on defense) and Denver (in orange, on offense) player on the field at 1, 2, 3, and 4 seconds after the ball snap is displayed, with Crosby highlighted in blue.
Note that the point size for each Las Vegas defender represents their STRAIN value at the different snapshots.

\begin{table}
\caption{Example of tracking data for a play during the Las Vegas Raiders versus Denver Broncos NFL game on October 17, 2021. The variables include frame identifier for each play (\textbf{frameId}); player position on the field (\textbf{x} and \textbf{y} coordinates); speed (\textbf{s}, in yards/second); acceleration (\textbf{a}, in yards/second$^2$); distance traveled from previous frame (\textbf{dis}, in yards); orientation (\textbf{o}, between 0 and 360 degrees); angle of motion (\textbf{dir}, between 0 and 360 degrees); and event annotation for each frame (\textbf{event}). The data shown here are for Raiders defensive end Maxx Crosby, and the frames included are between the ball snap and when the sack by Crosby is recorded. \label{tab:track}}
\centering
\begin{tabular}{ccccccccc}
\hline
\textbf{frameId} & \textbf{x} & \textbf{y} & \textbf{s} & \textbf{a} & \textbf{dis} & \textbf{o} & \textbf{dir} & \textbf{event} \\ 
\hline
7 & 67.68 & 29.89 & 0.34 & 1.57 & 0.04 & 124.86 & 88.21 & ball\_snap \\ 
8 & 67.76 & 29.89 & 0.69 & 2.13 & 0.08 & 124.07 & 89.59 & None \\ 
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots & \vdots \\ 
50 & 73.67 & 25.06 & 4.19 & 2.62 & 0.42 & 134.21 & 125.26 & qb\_sack \\ 
\hline
\end{tabular}
\end{table}

@@

Figure \ref{fig:pass_rushers_count} is a histogram of the number of pass rushers across every play in our data.
The pass rusher count ranges from 1 to 8, with 4 being the mode of the distribution.

@@

# Multilevel Model Without Team Effects

@@

We perform leave-one-week-out cross-validation to compare the multilevel models with and without team effects for play-level STRAIN.
For example, we train each model using data from weeks one through seven, and perform evaluation for week eight.
This procedure is repeated for every single holdout week, and we evaluate the model performance based on mean squared error (MSE).
Figure \ref{fig:fig_loocv_team_eff} shows the week-by-week holdout MSE for the aforementioned multilevel models.
We clearly see that a smaller MSE is achieved for each left-out week when accounting for team effects.
Furthermore, the model with team effects yields a lower overall holdout MSE ($4.1882$) than the one without random intercepts for teams ($4.4138$).
The out-of-sample performance justifies the inclusion of team effects in our proposed multilevel model in the paper.
